---
title: ""
output: md_document
date: ""
---

## Overview

Simplified version of a tomato growth and yield model for Tomato Trials <link to come>.

## References

 + [Jones, J. W., Dayan, E., Allen, L. H., Van Keulen, H., & Challa, H. (1991). A dynamic tomato growth and yield model (TOMGRO). Transactions of the ASAE, 34(2), 663-0672.](https://www.researchgate.net/profile/Leon-Allen-3/publication/40142875_A_dynamic_tomato_growth_and_yield_model_TOMGRO/links/58911ee9a6fdcc1b41453859/A-dynamic-tomato-growth-and-yield-model-TOMGRO.pdf)
 + [Jones, J. W., Kenig, A., & Vallejos, C. E. (1999). Reduced state–variable tomato growth model. Transactions of the ASAE, 42(1), 255-265.](https://www.researchgate.net/profile/James-Jones-30/publication/274529819_Reduced_state-variable_tomato_growth_model/links/5f3ed116a6fdcccc43db9fb2/Reduced-state-variable-tomato-growth-model.pdf)
 + [GitHub '@gyosit` gist of TOMGRO model in python](https://gist.github.com/gyosit/abeab4e595d7ddcd65b55c1270d240c8)
 

## Description

The file `tomgro.r` contains the required functions to simulate tomato yield based on daily average temperatures ($^\circ$C), photosynthetic photon flux density, PPFD (light intensity), 

## Tomato Trials examples


```{r}
source("tomgro.r")
## plant_type and percent_manure are user input 7x7 arrays
## where the path indecies (5, 1), (5, 2), (5, 3), (4, 5), (4, 6), (4, 7)
## are ignored (maybe set to "path" or NA if numeric)
## NOTE
## if only a single type of plant is used then there are only 30 seedlings
## and the function will pick up on this by IDing NA percent_manure values in the array
## example
plant_type <- matrix(sample(c("cherry", "heirloom"), replace = TRUE, size = 49), nrow = 7)
percent_manure <- matrix(sample(c(0, 10, 30, 50, 80), replace = TRUE, size = 49), nrow = 7)
## path index
idx <- cbind(c(5, 5, 5, 4, 4, 4,1:7), c(1:3, 5:7, rep(4,7)))
plant_type[idx] <-  "path"
plant_type
percent_manure[idx] <- NA
percent_manure
```

### Simulation

```{r}
## map recipe names to % values
recipes <- c("Control" = 0, "Recipe 1" = 10, "Recipe 2" = 30, "Recipe 3" = 50, "Recipe 4" = 80)
results <- tomgro(plant_type = plant_type, percent_manure = percent_manure, recipes = recipes, prob_die = c(0.1, 0.5))
results


```


### The 'TRUE' effect of manure

```{r, echo = FALSE, warning = FALSE,message = FALSE, out.width="100%"}
perc <- 0:100
func <- function(x, type = "cherry"){
    if(type == "cherry"){
    res = (3/5)*x - (1/175)*x^2 
    }else{
      if(type == "heirloom"){
    res = 4*sqrt(x) - (1/3)*x
      }
    }
  return(res)
}
data <- data.frame(Percentage = rep(perc, 2), value = c(func(perc), func(perc, "heirloom")), Type = rep(c("Cherry", "Heirloom"), each = length(perc)))
require(ggplot2)
ggplot(data, aes(x = Percentage, y = value, col = Type)) +
  geom_line(linewidth = 3) + theme_classic() + xlab("Percentage Manure") +
   scale_color_brewer(palette = "Set1") +
  ylab("Effect size (%)") +
  geom_vline(xintercept = range(perc[which(func(perc) > 15.5)]), col = "red", alpha = 0.5) +
  geom_vline(xintercept = range(perc[which(func(perc) > 13)]), col = "red", alpha = 0.5, lty = 2)+
  geom_vline(xintercept = range(perc[which(func(perc, "heirloom") > 11.5)]), col = "darkblue", alpha = 0.5) +
  geom_vline(xintercept = range(perc[which(func(perc, "heirloom") > 10)]), col = "darkblue", alpha = 0.5, lty = 2) +
  labs(subtitle = "Vertical lines indicate farm status if final answer within that range: thrived (solid) and survived (dashed), else mortgagee sale")
## cherry thrived perc[which(func(perc) > 15.5)]
## cherry survived perc[which(func(perc) > 13)]
## heirloom thrived perc[which(func(perc, "heirloom") > 11.5)]
## heirloom survived perc[which(func(perc, "heirloom") > 10)]
```

## Further examples & student plotting code


### Just 30 Cherry seedlings

```{r, warning = FALSE,message = FALSE, out.width="100%"}
## Just 30 Cherry tomato seedlings
## setting up data
plant_type <- matrix(rep("cherry", 49), nrow = 7) 
percent_manure <- matrix(sample(c(0, 10, 30, 50, 80), replace = TRUE, size = 49), nrow = 7)
## path index
idx <- cbind(c(5, 5, 5, 4, 4, 4,1:7), c(1:3, 5:7, rep(4,7)))
## six missing planters
idx_empty <- cbind(c(1:3, 5:7), c(1:3, rep(7, 3)))
plant_type[idx] <-  "path"
plant_type[idx_empty] <-  "empty"
percent_manure[rbind(idx, idx_empty)] <- NA
recipes <- c("Control" = 0, "Recipe 1" = 10, "Recipe 2" = 30, "Recipe 3" = 50, "Recipe 4" = 80)
results <- tomgro(plant_type = plant_type, percent_manure = percent_manure, recipes = recipes, prob_die = c(0.1, 0.5))
```


```{r, warning = FALSE,message = FALSE, out.width="100%"}
##plotting code for students
library(ggplot2)
ggplot(data = results, aes(x = recipe, y = fruit_dry_weight)) +
  geom_violin() + geom_dotplot(binaxis = "y", method = "histodot") +
  theme_bw() + xlab("Fertilizer Recipie") + ylab("Fruit Dry Weight (g/m^2)")
ggplot(data = results, aes(x = recipe, y = mature_fruit_weight)) +
  geom_violin() + geom_dotplot(binaxis = "y", method = "histodot") +
  theme_bw() + xlab("Fertilizer Recipie") + ylab("Mature Fruit Weight (g/m^2)")

```


### Just 30 Heirloom seedings

```{r, warning = FALSE,message = FALSE, out.width="100%"}
## Just 30 Cherry tomato seedlings
## setting up data
plant_type <- matrix(rep("heirloom", 49), nrow = 7) 
percent_manure <- matrix(sample(c(0, 10, 30, 60, 80), replace = TRUE, size = 49), nrow = 7)
## path index
idx <- cbind(c(5, 5, 5, 4, 4, 4,1:7), c(1:3, 5:7, rep(4,7)))
## six missing planters
idx_empty <- cbind(c(1:3, 5:7), c(1:3, rep(7, 3)))
plant_type[idx] <-  "path"
plant_type[idx_empty] <-  "empty"
percent_manure[rbind(idx, idx_empty)] <- NA
recipes <- c("Control" = 0, "Recipe 1" = 10, "Recipe 2" = 30, "Recipe 3" = 60, "Recipe 4" = 80)
results <- tomgro(plant_type = plant_type, percent_manure = percent_manure, recipes = recipes, prob_die = c(0.1, 0.5))
```

```{r,warning = FALSE,message = FALSE, out.width="100%"}
## plotting code for students
library(ggplot2)
## one tomato type
ggplot(data = results, aes(x = recipe, y = fruit_dry_weight)) +
  geom_violin() + geom_dotplot(binaxis = "y", method = "histodot") +
  theme_bw() + xlab("Fertilizer Recipie") + ylab("Fruit Dry Weight (g/m^2)")
ggplot(data = results, aes(x = recipe, y = mature_fruit_weight)) +
  geom_violin() + geom_dotplot(binaxis = "y", method = "histodot") +
  theme_bw() + xlab("Fertilizer Recipie") + ylab("Mature Fruit Weight (g/m^2)")

```

### Heirloom & Cherry seedings


```{r,warning = FALSE,message = FALSE, out.width="100%"}
plant_type <- matrix(sample(c("cherry", "heirloom"), replace = TRUE, size = 49), nrow = 7)
percent_manure <- matrix(sample(c(0, 10, 40, 60, 90), replace = TRUE, size = 49), nrow = 7)
## path index
idx <- cbind(c(5, 5, 5, 4, 4, 4,1:7), c(1:3, 5:7, rep(4,7)))
plant_type[idx] <-  "path"
percent_manure[idx] <- NA
recipes <- c("Control" = 0, "Recipe 1" = 10, "Recipe 2" = 40, "Recipe 3" = 60, "Recipe 4" = 90)
results <- tomgro(plant_type = plant_type, percent_manure = percent_manure, recipes = recipes, prob_die = c(0.1, 0.5))
```

```{r, warning = FALSE,message = FALSE, out.width="100%"}
## plotting code for students
library(ggplot2)
ggplot(data = results, aes(x = recipe, y = fruit_dry_weight, fill = plant)) +
  geom_violin() + geom_dotplot(binaxis = "y", method = "histodot") +
  theme_bw() + xlab("Fertilizer Recipie") + ylab("Fruit Dry Weight (g/m^2)") +
  facet_wrap(~plant)
```