---
title: ""
output: md_document
date: ""
---

## Overview

Simplified version of a tomato growth and yield model for Tomato Trials <link to come>.

## References

 + [Jones, J. W., Dayan, E., Allen, L. H., Van Keulen, H., & Challa, H. (1991). A dynamic tomato growth and yield model (TOMGRO). Transactions of the ASAE, 34(2), 663-0672.](https://www.researchgate.net/profile/Leon-Allen-3/publication/40142875_A_dynamic_tomato_growth_and_yield_model_TOMGRO/links/58911ee9a6fdcc1b41453859/A-dynamic-tomato-growth-and-yield-model-TOMGRO.pdf)
 + [Jones, J. W., Kenig, A., & Vallejos, C. E. (1999). Reduced state–variable tomato growth model. Transactions of the ASAE, 42(1), 255-265.](https://www.researchgate.net/profile/James-Jones-30/publication/274529819_Reduced_state-variable_tomato_growth_model/links/5f3ed116a6fdcccc43db9fb2/Reduced-state-variable-tomato-growth-model.pdf)
 + [GitHub '@gyosit` gist of TOMGRO model in python](https://gist.github.com/gyosit/abeab4e595d7ddcd65b55c1270d240c8)
 

## Description

The file `tomgro.r` contains the required equations to simulate tomato yield based on daily average temperatures ($^\circ$C) and photosynthetic photon flux density, PPFD (light intensity).

## Tomato Trials example

```{r}
source("tomgro.r")
## set up environmental data
require(lubridate)
dates <- seq(lubridate::ymd_h("2024-01-01-00"), lubridate::ymd_h("2024-06-30-23"), by = "days")
## from left to right PPFD gets stronger
PPFD_gradient <- matrix(rep(seq(140, 200, length.out = 7), each = 7), nrow = 7)
## from top to bottom temp gets cooler
temp_gradient <- matrix(rep(seq(28, 21, length.out = 7), each = 7), nrow = 7, byrow = TRUE)
## path index
idx <- cbind(c(5, 5, 5, 4, 4, 4,1:7), c(1:3, 5:7, rep(4,7)))
temp_gradient[idx] <- PPFD_gradient[idx] <-  NA
```

### Baseline Greenhouse environmenlat variables

```{r, echo = FALSE, warning = FALSE}
library(reshape2)
library(ggplot2)

longData <- melt(temp_gradient)
longData<-longData[longData$value!=0,]

t <- ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_tile(aes(fill=value), col = "black") + 
  scale_fill_gradient(low="grey90", high="red") +
  theme_void() + theme(legend.position = "none") + ggtitle("Temperature gradient")

longData <- melt(PPFD_gradient)
longData<-longData[longData$value!=0,]

p <- ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_tile(aes(fill=value), col = "black") + 
  scale_fill_gradient(low="grey90", high="red") +
  theme_void() + theme(legend.position = "none") + ggtitle("PPFD gradient")

patchwork::wrap_plots(t, p)
```


### Simulating data based on environmental data alone

Note plant and fertilizer choices TODO (just involves changing coefficients of growth equations)

```{r}
## output matricies
outputW <- outputWf <- outputWm <- matrix(NA, nrow = 7, ncol = 7)
for(i in 1:7){ ## rows
    for(j in 1:7){ ## cols
        if(!is.na(temp_gradient[i, j]) & !is.na(PPFD_gradient[i, j])){
            temp <- rnorm(length(dates), temp_gradient[i, j], 1) +
                (cos(lubridate::day(dates))/(length(lubridate::day(dates))/5))
            ppfd <- rnorm(length(dates), PPFD_gradient[i, j], 20)
            gr <- grow(T = temp, PPFD = ppfd)
            outputW[i, j] <- gr[1]
            outputWf[i, j] <- gr[2]
            outputWm[i, j] <- gr[3]
        }
    }
}
## Structure 
output <- data.frame(row = rep(1:7, times = 7),
                     col = rep(1:7, each = 7),
                     total_plant_weight = c(outputW),
                     fruit_dry_weight = c(outputWf),
                     mature_fruit_weight = c(outputWm))
output <- subset(output, !is.na(output$total_plant_weight))
## structure of data to download
head(output)
```

#### Illustration of output

```{r, echo = FALSE}
ggplot(output, aes(y = row, x = col)) + 
  geom_tile(aes(fill=mature_fruit_weight), col = "black") + 
  scale_fill_gradient(low="grey90", high="red") +
  theme_void() + theme(legend.position = "none") + ggtitle("Mature Fruit Dry Weight")


```